---
title: "The E-runs"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Robin found a weird pattern in his simulations in which the variance in the
number of cross-cohort half-sibling pairs was greater than the mean.  This is,
perhaps, to be expected when population sizes are quite small relative to the sample
sizes so that there are many pairs of sibs from the same parent (or, somewhat equivalently,
there are many sib pairs that share members---i.e. they are part of the same connected
component if you make a graph).

When the population size gets bigger, and thus the sampling is getting sparser, in some sense,
it seems that that effect should attenuate, and the variance in the number of half-sibling pairs
should get closer to the number of mean.

The E-runs are a series of simulations in which I simulate using Robin's basic demography of fish
that live for 10 years.  To see the sample sizes (per year) and the downsampled sizes (over all 5 years
of sampling) that we are dealing with we can look at the snakefile:
```{r}
system("head -n 8 Snakefile-E-runs")
```


This is set up to do 200 reps of each scenario in parallel on the cluster, and it should never
eat up more than 25 nodes at a time.  The total adult population size is roughly twice the cohort
size.  The true pop size value gets recorded.

To launch the runs, we do like this:
```sh

```
